---
title: "HW 03 - Great Lakes Fish Datasets"
author: "Vincent Baker"
date: "2/27/23"
output: html_document
---

## Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse)
```

```{r load-data, message = FALSE}
catch_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-08/fishing.csv')
catch_data

fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
fish_stock
```

## Exercises

### Exercise 1

```{r catch-data-subset}
catch_data <- select(catch_data, year, lake, species, region, values)
catch_data
```

```{r fish-stock-subset}
fish_stock <- select(fish_stock, YEAR, MONTH, LAKE, STATE_PROV, SPECIES, WEIGHT)
fish_stock
```


### Exercise 2

```{r glimpse-catch-data}
glimpse(catch_data)
```

```{r fish-stock}
glimpse(fish_stock)
```


### Exercise 3

```{r filter-catch, eval = TRUE}
catch_data <- catch_data %>% 
  filter(region == "U.S. Total",
         species == "Chinook Salmon",
         lake == "Michigan")
```

```{r filter-stock, eval = TRUE}
fish_stock <- fish_stock %>% 
  filter(STATE_PROV %in% c("MI", "WI", "IL", "IN"),
         SPECIES == "CHS",
         LAKE == "MI")
```


### Exercise 4

There are `r nrow(catch_data)` in this dataset.
The first year is `r min(catch_data$year)` in this dataset.
The last year is `r max(catch_data$year)` in this dataset.

There are `r nrow(fish_stock)` in this dataset.
The first year is `r min(fish_stock$YEAR)` in this dataset.
The last year is `r max(fish_stock$YEAR)` in this dataset.

### Exercise 5

```{r fish-stock-yearly}
fish_stock_yearly <- fish_stock %>%
  group_by(YEAR) %>%
  summarize(STOCK_WEIGHT_YR = sum(WEIGHT, na.rm = TRUE))
```


### Code chunks provided

Remember to change eval = FALSE, to eval = TRUE, when you are ready to run the code/knit the document.

```{r finalize-fish-stock, eval = TRUE}
fish_stock_yearly <- fish_stock_yearly %>% 
  rename(year = YEAR) %>% 
  mutate(stock_weight_yr_lbs = as.integer(STOCK_WEIGHT_YR * 0.454 / 1000)) %>% # 1 lb = 1 kg * 0.454
  select(year, stock_weight_yr_lbs)
```

```{r finalize-catch-data, eval = TRUE}
catch_data <- catch_data %>% 
  rename(catch_weight_yr_lbs = values)
```

### Exercise 6

```{r left-join}
left_join(fish_stock_yearly, catch_data)
```

All data from fish_stock_yearly and catch_data was kept.

```{r right-join}
right_join(fish_stock_yearly, catch_data)
```

All of the variables from fish_stock_yearly were kept, but only observations that included data from the catch_data data frame.

```{r inner-join}
inner_join(fish_stock_yearly, catch_data)
```

More of the same from the previous table.

```{r full-join}
full_join(fish_stock_yearly, catch_data)
```

All observations from the fish_stock_yearly and catch_data were kept.

**For these last two examples we provide the code, just write narrative for both**

```{r anti-join-v1, eval = TRUE}
catch_data %>% 
  anti_join(fish_stock_yearly)
```

All rows were lost and all variables were kept except 'fish_stock_yr_ibs.'

```{r anti-join-v2, eval = TRUE}
fish_stock_yearly %>% 
  anti_join(catch_data)
```

All variables from fish_stock_yearly were kept and all observations except those from 1992-1999 were kept.


### Exercise 7

```{r pivot-longer, eval = TRUE}
catch_data %>% 
  left_join(fish_stock_yearly) %>% 
  pivot_longer(names_to = "activity",
               values_to = "weight_lbs",
               cols = 5:6)
```


### Exercise 8 -.5

```{r final-visualization, eval = TRUE}
catch_data %>% 
  left_join(fish_stock_yearly) %>% 
  pivot_longer(names_to = "activity",
               values_to = "weight_lbs",
               cols = 5:6) %>%
  ggplot(aes(x = year, y = weight_lbs, color = activity)) +
  geom_line() +
  theme_bw()
  
```

The size of the fish catch has rapidly increased since the early 90s to 1995, then sharply decreased towards 2000 and has shown a steady decline in size; the amount of fish in Lake Michigan steadily increased toward the late 90s, then has steadily decreased over time. 
