---
title: "Lab 03 - Great Lakes Fish Stocking"
author: "Vincent Baker"
date: "2/21/23"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

There are 56,232 observations and 8 variables in this dataset. 

```{r fish-stock}
glimpse(fish_stock)
```


### Exercise 2

```{r fish-species}
salmon_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("ATS", "CHS", "COS"),
    !is.na(WEIGHT),
     LAKE != "ON"
  )
```

```{r salmon-stock}
glimpse(salmon_stock)
```

There are 5,131 observations and 8 variables in the dataset. Lake and state province are character strings and the rest are numerical data. 

### Exercise 3

Remove this text, and add your answer for Exercise 3 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.
Add code chunks as needed.
Don't forget to label your code chunk.
Do not use spaces in code chunk labels.
