---
title: "Lab 03 - Great Lakes Fish Stocking"
author: "Vincent Baker"
date: "2/21/23"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
fish_stock <- read_csv("data/great-lakes-fish-stock.csv")
```

## Exercises

### Exercise 1

There are 56,232 observations and 8 variables in this dataset. 

```{r fish-stock}
glimpse(fish_stock)
```


### Exercise 2

```{r fish-species}
salmon_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("ATS", "CHS", "COS"),
    !is.na(WEIGHT),
     LAKE != "ON"
  )
```

```{r salmon-stock}
glimpse(salmon_stock)
```

There are 5,131 observations and 8 variables in the dataset. Lake and state province are character strings and the rest are numerical data. 

### Exercise 3

```{r salmon-weight}
salmon_stock <- salmon_stock %>%
  mutate(
    AVERAGE = WEIGHT / NO_STOCKED * 1000
  )
salmon_stock
```

mutate() adds variables and keeps the existing ones. For example, it combined 'WEIGHT' and 'NO_STOCKED', and it created AVERAGE (the average weight for each salmon in each stocking event in grams).


### Exercise 4

```{r summary-statistics}
salmon_stock %>% 
  group_by(LAKE) %>% 
  filter(YEAR > 1999) %>% 
  summarize(MEAN_NO_STOCKED = mean(NO_STOCKED, na.rm = TRUE),
            MEAN_WEIGHT = mean(WEIGHT, na.rm = TRUE)) %>% 
  arrange(desc(MEAN_NO_STOCKED))
```

Lake Superior has the largest mean salmon per stocking event and Lake Michigan has the largest mean weight per stocking event.

### Exercise 5

```{r fish-species, weight, summary-statistics}
trout_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("BKT", "BKN", "LAT", "RBT", "TRT"),
    !is.na(WEIGHT),
    LAKE != "ON"
  )

trout_stock <- trout_stock %>%
  mutate(
    AVERAGE = WEIGHT / NO_STOCKED * 1000)
    trout_stock
    
trout_stock %>% 
  group_by(LAKE) %>% 
  summarize(MEAN_NO_STOCKED = mean(NO_STOCKED, na.rm = TRUE),
            MEAN_WEIGHT = mean(WEIGHT, na.rm = TRUE)) %>% 
  arrange(desc(MEAN_NO_STOCKED))
```

This code chunk above shows the data for trout species similar to the salmon species dataset. 

### Exercise 6

```{r annual-plot}
trout_and_salmon_stock <- fish_stock %>%
  filter(
    SPECIES %in% c("CHS", "BKT"),
    !is.na(WEIGHT)) %>%
    group_by(YEAR, SPECIES, LAKE) %>%
      summarise(ANNUAL_NO_STOCKED = sum(NO_STOCKED, na.rm = TRUE))

ggplot(trout_and_salmon_stock, aes(x = YEAR, y = ANNUAL_NO_STOCKED, color = SPECIES)) +
  geom_line() +
    facet_wrap(~LAKE, ncol = 2) +
      theme_bw() +
      labs(
        x = "YEAR",
        y = "Annual number of salmon and trout stocked",
        color = "Species"
      )
```

Chinook salmon is stocked at a higher rate.
Lake Michigan had the largest fish stocking activities. 
Stocking reached its peak between 1988-2000 (~1989).
